A Type System for Safe and Scalable Parallelism using Logical Regions \\ (Extended Version)

### 2012-07-19 doc/popl2013full/langdef.tex
edits

\noindent Two additional properties are proven for each subexpression:
\begin{itemize}
\item $\valueconsist{v}{T}$, the expression yields a value of the right type
\item $\traceconsist{E}$, values written to the store match the heap types of their locations
\end{itemize}
Figure~\ref{fig:constprop} gives the precise definition of these
properties.\footnote{Note that in the definition of
$\mapconsist{\Omega}$, the quantified variables $r_i$ and $r_j$
range over both logical and concrete regions.  Since $M$'s domain is
logical regions, we extend $M$ to constraints with mixed logical and
concrete regions in the standard way by defining
$M \llbracket \rho \rrbracket = \rho$.}

\begin{figure*}
\centering{\footnotesize
$
\begin{array}{l}
\begin{array}{ll}
\begin{array}{lll}
\mapconsist{\Omega} & \Leftrightarrow & (\forall r_i, r_j. r_i \leq r_j \in \Omega \Rightarrow M(r_i) \subseteq M(r_j))\ \wedge \\
              &                 & (\forall r_i, r_j. r_i * r_j \in \Omega \Rightarrow M(r_i) \cap M(r_j) = \emptyset) \\
\localconsist{L}{\Gamma} & \Leftrightarrow & \forall (id,v) \in L. \valueconsist[]{v}{M \llbracket \Gamma \rrbracket (id)} \\
\storeconsist{S}{H} & \Leftrightarrow & \forall (l,v) \in S. \valueconsist[]{v}{H(l)} \\
\end{array} & 
\begin{array}{lll}
\traceconsist{E} & \Leftrightarrow & (\forall l, c, v. write(l,c, v) \in E \Rightarrow \valueconsist[]{v}{H(l)})\ \wedge \\
& & (\forall id, l, v. reduce_{id}(l, v) \in E \Rightarrow \\
& & (M \llbracket \Gamma \rrbracket (id) = (\hat T_1, \hat T_2), \emptyset, \emptyset \rightarrow \hat T_1) \wedge H(l) = \hat T_1 \wedge \valueconsist[]{v}{\hat T_2}) \\
\end{array}
\end{array} \vspace{3mm} \\
\begin{array}{l@{\hspace{1cm}}l}
\begin{array}{l}
\valueconsist[]{bv}{bool} \\
\valueconsist[]{iv}{int} \\
\valueconsist[]{null}{\hat T@\rho} \\
\end{array} &
\begin{array}{lll}
\valueconsist[]{l}{\hat T@\rho} & \Leftrightarrow & l \in \rho \wedge H(l) = \hat T \\
\valueconsist[]{\tuple{v_1}{v_2}}{\tuple{\hat T_1}{\hat T_2}} & \Leftrightarrow & (\valueconsist[]{v_1}{\hat T_1}) \wedge (\valueconsist[]{v_2}{\hat T_2}) \\
\valueconsist[]{\langle \langle \oton{\rho}, v \rangle \rangle}{\text{rr}[\oton{r}]\ \hat T\text{ where }\Omega} & \Leftrightarrow & (\valueconsist[]{v}{T[\rho_1/r_1, \ldots \rho_n/r_n]}) \wedge (\mapconsist[\{ (r_i, \rho_i) \}]{\Omega}) \\
\valueconsist[]{K}{coloring(\rho)} & \Leftrightarrow & \forall l_1, v_1. (l_1, v_1) \in K \Rightarrow (l_1 \in \rho\ \wedge \\
& & \forall l_2, v_2. (l_2, v_2) \in K \Rightarrow (l_1 \not= l_2) \vee (v_1 = v_2)) \\
\end{array}
\end{array}
\end{array}
$
}
\caption{Consistency Properties}
\label{fig:constprop}
\vspace{-5mm}
\end{figure*}

@@

The proof is a standard induction on the structure of the derivation
For each of 
the Core Legion expressions there are six separate parts to the proof,
one for each of the three consistency properties that must be shown
for the execution environment of subexpressions and one for each of
the three properties of the execution's results.
Many of the cases are similar, and benefit from
the use of the following straighforward lemmas:
\begin{lem}
\label{lemma:heapconst:apply}
\rm
If $\storeconsist{S}{H}$ and $\traceconsist{E}$, then $\storeconsist{\text{apply}(S, E)}{H}$.
\end{lem}

@@

We first give an outline of the idea behindthe proof.  For each subexpression of a compound expression,
there are three consistency properties we must prove to be able to apply induction: \\

@@

\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\storeconsist{S}{H}$ -  The heap typing consistency of all stores used in subexpressions follows directly from the following
lemma: If $\storeconsist{S}{H}$ and $\traceconsist{E}$, then $\storeconsist{\text{apply}(S, E)}{H}$.

@@

\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\traceconsist{E}$ -  The type consistency of the values in an expression's memory
trace follows from the following lemma applied to the traces of subexpressions:
if $\traceconsist{E_1}$ and $\traceconsist{E_2}$ and $any\_interleave(E', E_1, E_2)$, then $\traceconsist{E'}$.
New memory operations are added by {\tt write} and {\tt reduce} expressions, but consistency follows
directly from the induction hypothesis.  Finally, the consistency of the
values in a called function's memory trace is addressed in the same way as the return value.

@@

\begin{proof}
The proof of the theorem is by cases on the structure of the derivation of $e$.
\begin{itemize}

\item
Case [E-Bool]. By assumption, 
$$\opsenv{bv}{bv}{[]}$$ and
$$\typeenv{bv}{T}$$ and
$\mapconsist{\Omega}$ and
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The only type rule that can apply is [T-Bool], so $T = bool$.
Therefore $\valueconsist{bv}{bool}$ and 
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$ all follow trivially.

@@

\item Case [E-Null].
By assumption, 
$$\opsenv{\text{null\ }T@r}{null}{[]}$$ and 
$$\typeenv{\text{null}}{T'}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
The only type rule that can apply is [T-Null], so $T' = T@r$.
Therefore $\valueconsist{null}{T@r}$, $\traceconsist{[]}$ and $\privconsist{[]}{\Phi}$ all follow trivially.

@@

\item Case [E-IfElse-T].
 By assumption, 
$$\opsenv{\text{ if } e_1 \text{ then } e_2 \text{ else } e_3}{v_2}{E}$$ and 
$$\typeenv{\text{ if } e_1 \text{ then } e_2 \text{ else } e_3}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.

For the subexpression $e_1$, 
from [E-IfElse-T] we know $$\opsenv{e_1}{true}{E_1 \dplus E_2}$$
From [T-IfElse], we know that $$\typeenv{e_1}{bool}$$
Clearly $\valueconsist{true}{bool}$, and by induction we have  $\traceconsist{E_1}$ and 
      $\privconsist{E_1}{\Phi}$.

@@

\item Case [E-Newcolor].
By assumption, 
$$\opsenv{\text{newcolor } r}{K}{[]}$$ 
where  $K = \{ (l_1,v_1),\ldots,(l_p,i_p) \}$ such that
$l_i \in M(r)$ for all $i$ and $l_1,\ldots,l_p$ are pairwise distinct locations.
In addition, we have the assumptions
$$\typeenv{\text{newcolor } r}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The only type rule that can apply if [T-Newcolor], so can infer that $T = \text{coloring}(r)$.
From the structure of $K$ it follows that $\valueconsist{K}{\text{coloring(r)}}$.
The other two conclusions,
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$, follow trivially.

\item  Case [E-Color].
By assumption, 
$$\opsenv{\text{color($e_1$, $e_2$, $e_3$)}}{K'}{E'}$$ and
$$\typeenv{\text{color($e_1$, $e_2$,$e_3$)}}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The only type rule that can apply if [T-Color], so we know $T = \text{coloring}(r)$.

@@

For the subexpression $e_3$, 
from [E-Color] we know $$\opsenv[S=S'']{e_3}{v}{E_3}$$ where $S'' \text{apply}(S',E_2)$.
From [T-Color], we know that $$\typeenv{e_3}{int}$$
By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S''}{H}$.
By induction, $\valueconsist{v}{int}$ and $\traceconsist{E_3}$ and 
      $\privconsist{E_3}{\Phi}$.

Now $K'$ is built from $K$ by extending $K$ with a mapping for location $l$.  First, because $\valueconsist{l}{T'@r}$ and
$\valueconsist{K}{\text{coloring(r))}$ and because $K'$ excludes any previous mapping for $l$ (guaranteeing the the coloring is a function), we conclude $\valueconsist{K'}{\text{coloring}(r)}}$.  From [E-Color], we also know
$$valid\_interleave(S, C, E', E_1, E_2, E_3)$$ and by applying Lemma~\ref{lemma:heapconst:effects2} twice
we can conclude $\traceconsist{E'}$.  Similarly, by applying Lemma~\ref{lemma:effsound:effects2} twice
we can conlude $\privconsist{E'}{\Phi}$.

\item  Case [E-UpRgn].
By assumption, $$\opsenv{\text{upregion}(e,\oton{r})}{v}{E}$$ and
$$\typeenv{\text{upgregion}(e,\oton{r})}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The type rule must be [T-UpRgn], so we also know $T = T' @ (\oton{r})$.

@@

To show $\valueconsist{v}{T'@(\oton{r})}$, we first observe that from
$\valueconsist{v}{T'@(r_1',\ldots,r_k')}$ we know
there is an $i$ such that $v \in M(r_i')$ and 
$H(v) = \hat T'$.   From [T-UpRgn]
we know $\forall i. \exists j. r_i' \leq r_j \in \Omega^*$.  Combining this with
$\mapconsist{\Omega}$ and Lemma~\ref{lemma:omegaclosure}, we know there
is some $r_j$ such that $v \in M(r_j)$.
Therefore $\valueconsist{v}{T'@(\oton{r})}$.

\item Case [E-DownRgn].
By assumption, $$\opsenv{\text{downregion}(e,\oton{r})}{v}{E}$$ and
$$\typeenv{\text{downgregion}(e,\oton{r})}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The type rule must be [T-DownRgn], so we also know $T = T' @ (\oton{r})$.

@@

To show $\valueconsist{v}{T}$ there are two cases.
If $v = M(l)$ then the claim follows from $\valueconsist{l}{T@(\oton{r})}$.
If $v$ is instead some $v'$ of type $H(l)$ the result also follows because $H(l) = \hat{T}$.
Next, $\traceconsist{E \dplus [ read(l,excl,v)]}$ follows immediately from $\traceconsist{E}$ as
any read operation is consistent (the resriction is only on writes and reductions).

@@

Observe that 
$$M \llbracket \exists r'_1, \ldots r'_n.\ T_2\text{ where }\Omega_1 \rrbracket$$
replaces all free region names in $T_2$ and particularly $\Omega_1$ by concrete regions, so
that the only region names remaining are $r'_1,\ldots,r'_n$.  It is then easy to show that
$$\mapconsist[\{(r'_i,\rho_i)\}]{M \backslash \{r'_1,\ldots, r'_n \}\llbracket \Omega_1 \rrbracket}$$
from which $\valueconsist{v'}{T_1}$ follows (see Figure~\ref{fig:constprop}).

@@

To show $\localconsist[M']{L'}{\Gamma'}$, we must show
$\valueconsist[M']{L'(x)}{\Gamma'(x)}$ for all identifiers $x$ in
the domain of $\Gamma'$.  For any $id' \neq id$, the statement is equivalent
to showing $\valueconsist[M']{L(x)}{\Gamma(x)}$, which must hold becuase $M'$ does 
does not change any existing binding in $M$.  For $id$, we have
$\valueconsist[M']{v_1}{T_1[r_1/r'_1,\ldots,r_k/r'_k]}$, which is implied by
$\valueconsist{\langle \langle \otok{k}{\rho},v \rangle\rangle}{T_1}$.

@@

To show that
$\traceconsist{E_{n+1}}$ under $M'$ implies $\traceconsist{E_{n+1}}$
under $M$, note that $H(\cdot)$ does not depend on the mapping (because it includes only concrete, not logical, regions) so the
types of locations are unaffected by the mapping (see
Figure~\ref{fig:constprop}).  Now, the environment
$M' \llbracket \Gamma' \rrbracket$ is used to look up reduction
functions, but the reduction function signatures are restricted to
have no regions named in their arguments and so
$$M \llbracket \Gamma' \rrbracket(id)  = M' \llbracket \Gamma' \rrbracket(id)$$
Finally, from 
$$valid\_interleave(S, C, E'', E', E_{n+1})$$
and Lemma~\ref{lemma:heapconst:effects2} we conclude $\traceconsist{E''}$.

For the last consistency property of the function call, we first observe
that $\privconsist[M']{E_{n+1}}{\Phi'}$ is equivalent to 
$$\privconsist{E_{n+1}}{\Phi[r_1/r'_1,\ldots,r_k/r'_k]}$$
because, again, $E_{n+1}$ does not mention logical region names and
$M' \llbracket r' \rrbracket = M \llbracket r \rrbracket$.
From [T-Call] we have 
$$\Phi[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Phi^*$$
and so
$\privconsist{E_{n+1}}{\Phi^*}$.  Applying Lemma~\ref{lemma:phiclosure} we conclude
that $\privconsist{E_{n+1}}{\Phi}$.
Finally, from Lemma~\ref{lemma:effsound:effects2} we conclude $\privconsist{E''}{\Phi}$.
\end{itemize}
\end{proof}

>>>>

\noindent Two additional properties are proven for each subexpression:
\begin{itemize}
\item $\valueconsist{v}{T}$, the expression yields a value of the right type
\item $\traceconsist{E}$, values written to the store match the heap types of their locations
\end{itemize}
Figure~\ref{fig:constprop} gives the precise definition of these
properties.\footnote{Note that in the definition of
$\mapconsist{\Omega}$, the quantified variables $r_i$ and $r_j$
range over both logical and physical regions.  Since $M$'s domain is
logical regions, we extend $M$ to constraints with mixed logical and
physical regions in the standard way by defining
$M \llbracket \rho \rrbracket = \rho$.}

\begin{figure*}
\centering{\footnotesize
$
\begin{array}{l}
\begin{array}{ll}
\begin{array}{lll}
\mapconsist{\Omega} & \Leftrightarrow & (\forall r_i, r_j. r_i \leq r_j \in \Omega \Rightarrow M(r_i) \subseteq M(r_j))\ \wedge \\
              &                 & (\forall r_i, r_j. r_i * r_j \in \Omega \Rightarrow M(r_i) \cap M(r_j) = \emptyset) \\
\localconsist{L}{\Gamma} & \Leftrightarrow & \forall (id,v) \in L. \valueconsist[]{v}{M \llbracket \Gamma \rrbracket (id)} \\
\storeconsist{S}{H} & \Leftrightarrow & \forall (l,v) \in S. \valueconsist[]{v}{H(l)} \\
\end{array} & 
\begin{array}{lll}
\traceconsist{E} & \Leftrightarrow & (\forall l, c, v. write(l,c, v) \in E \Rightarrow \valueconsist[]{v}{H(l)})\ \wedge \\
& & (\forall id, l, v. reduce_{id}(l, v) \in E \Rightarrow \\
& & (M \llbracket \Gamma \rrbracket (id) = (\hat T_1, \hat T_2), \emptyset, \emptyset \rightarrow \hat T_1) \wedge H(l) = \hat T_1 \wedge \valueconsist[]{v}{\hat T_2}) \\
\end{array}
\end{array} \vspace{3mm} \\
\begin{array}{l@{\hspace{1cm}}l}
\begin{array}{l}
\valueconsist[]{bv}{bool} \\
\valueconsist[]{iv}{int} \\
\valueconsist[]{null}{\hat T@\rho} \\
\end{array} &
\begin{array}{lll}
\valueconsist[]{l}{\hat T@\rho} & \Leftrightarrow & l \in \rho \wedge H(l) = \hat T \\
\valueconsist[]{\tuple{v_1}{v_2}}{\tuple{\hat T_1}{\hat T_2}} & \Leftrightarrow & (\valueconsist[]{v_1}{\hat T_1}) \wedge (\valueconsist[]{v_2}{\hat T_2}) \\
\valueconsist[]{\langle \langle \oton{\rho}, v \rangle \rangle}{\text{rr}[\oton{r}]\ \hat T\text{ where }\hat \Omega} & \Leftrightarrow & (\valueconsist[]{v}{T[\rho_1/r_1, \ldots \rho_n/r_n]}) \wedge (\mapconsist[\{ (r_i, \rho_i) \}]{\hat \Omega}) \\
\valueconsist[]{K}{coloring(\rho)} & \Leftrightarrow & \forall l_1, v_1. (l_1, v_1) \in K \Rightarrow (l_1 \in \rho\ \wedge \\
& & \forall l_2, v_2. (l_2, v_2) \in K \Rightarrow (l_1 \not= l_2) \vee (v_1 = v_2)) \\
\end{array}
\end{array}
\end{array}
$
}
\caption{Consistency Properties}
\label{fig:constprop}
\vspace{-5mm}
\end{figure*}

@@

The proof is a standard induction on the structure of the derivation
For each of 
the Core Legion expressions there are six separate parts to the proof,
one for each of the three consistency properties that must be shown
for the execution environment of subexpressions and one for each of
the three properties of the execution's results.
Many of the cases are similar, and benefit from
the use of the following straightforward lemmas:
\begin{lem}
\label{lemma:heapconst:apply}
\rm
If $\storeconsist{S}{H}$ and $\traceconsist{E}$, then $\storeconsist{\text{apply}(S, E)}{H}$.
\end{lem}

@@

We first give an outline of the idea behind the proof.  For each subexpression of a compound expression,
there are three consistency properties we must prove to be able to apply induction: \\

@@

\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\storeconsist{S}{H}$ -  The heap typing consistency of all stores used in subexpressions follows directly from
Lemma~\ref{lem:heapconst:apply}.

@@

\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\traceconsist{E}$ -  The type consistency of the values in an expression's memory
trace follows from Lemma~\ref{lem:heapconst:effects1} and Lemma~\ref{lem:heapconst:effects2}.
New memory operations are added by {\tt write} and {\tt reduce} expressions, but consistency follows
directly from the induction hypothesis.  Finally, the consistency of the
values in a called function's memory trace is addressed in the same way as the return value.

@@

\begin{proof}
The proof of the theorem is by cases on the structure of the derivation of $e$.
{\begin{itemize}
\setlength{\itemsep}{4pt}
\item
Case [E-Bool]. By assumption, 
$$\opsenv{bv}{bv}{[]}$$ and
$$\typeenv{bv}{T}$$ and
$\mapconsist{\Omega}$ and
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The only type rule that can apply is [T-Bool], so $T = bool$.
Therefore $\valueconsist{bv}{bool}$ and 
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$ all follow trivially.

@@

\item Case [E-Null].
By assumption, 
$$\opsenv{\text{null\ }T@r}{null}{[]}$$ and 
$$\typeenv{\text{null\ }T@r}{T'}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
The only type rule that can apply is [T-Null], so $T' = T@r$.
Therefore $\valueconsist{null}{T@r}$, $\traceconsist{[]}$ and $\privconsist{[]}{\Phi}$ all follow trivially.

@@

\item Case [E-IfElse-T].
 By assumption, 
$$\opsenv{\text{ if } e_1 \text{ then } e_2 \text{ else } e_3}{v_2}{E_1 \dplus E_2}$$ and 
$$\typeenv{\text{ if } e_1 \text{ then } e_2 \text{ else } e_3}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.

For the subexpression $e_1$, 
from [E-IfElse-T] we know $$\opsenv{e_1}{true}{E_1}$$
From [T-IfElse], we know that $$\typeenv{e_1}{bool}$$
Clearly $\valueconsist{true}{bool}$, and by induction we have  $\traceconsist{E_1}$ and 
      $\privconsist{E_1}{\Phi}$.

@@

\item Case [E-Newcolor].
By assumption, 
$$\opsenv{\text{newcolor } r}{K}{[]}$$ 
where  $K = \{ (l_1,v_1),\ldots,(l_p,i_p) \}$ such that
$l_i \in M(r)$ for all $i$ and $l_1,\ldots,l_p$ are pairwise distinct locations.
In addition, we have the assumptions
$$\typeenv{\text{newcolor } r}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The only type rule that can apply is [T-Newcolor], so we can infer that $T = \text{coloring}(r)$.
From the structure of $K$ it follows that $\valueconsist{K}{\text{coloring(r)}}$.
The other two conclusions,
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$, follow trivially.

\item  Case [E-Color].
By assumption, 
$$\opsenv{\text{color($e_1$, $e_2$, $e_3$)}}{K'}{E'}$$ and
$$\typeenv{\text{color($e_1$, $e_2$,$e_3$)}}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The only type rule that can apply is [T-Color], so we know $T = \text{coloring}(r)$.

@@

For the subexpression $e_3$, 
from [E-Color] we know $$\opsenv[S=S'']{e_3}{v}{E_3}$$ where $S'' = \text{apply}(S',E_2)$.
From [T-Color], we know that $$\typeenv{e_3}{int}$$
By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S''}{H}$.
By induction, $\valueconsist{v}{int}$ and $\traceconsist{E_3}$ and 
      $\privconsist{E_3}{\Phi}$.

Now $K'$ is built from $K$ by extending $K$ with a mapping for location $l$.  First, because $\valueconsist{l}{T'@r}$ and
$\valueconsist{K}{\text{coloring}(r)}$ and because $K'$ excludes any previous mapping for $l$ (guaranteeing the the coloring is a function), we conclude $\valueconsist{K'}{\text{coloring}(r)}$.  From [E-Color], we also know
$$valid\_interleave(S, C, E', E_1, E_2, E_3)$$ and by applying Lemma~\ref{lemma:heapconst:effects2} twice
we can conclude $\traceconsist{E'}$.  Similarly, by applying Lemma~\ref{lemma:effsound:effects2} twice
we can conclude $\privconsist{E'}{\Phi}$.

\item  Case [E-UpRgn].
By assumption, $$\opsenv{\text{upregion}(e,\oton{r})}{v}{E}$$ and
$$\typeenv{\text{upregion}(e,\oton{r})}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The type rule must be [T-UpRgn], so we also know $T = T' @ (\oton{r})$.

@@

To show $\valueconsist{v}{T'@(\oton{r})}$, we first observe that from
$\valueconsist{v}{T'@(r_1',\ldots,r_k')}$ we know
there is an $i$ such that $v \in M(r_i')$ and 
$H(v) = \hat T'$.   From [T-UpRgn]
we know $\forall i. \exists j. r_i' \leq r_j \in \Omega^*$.  Combining this with
$\mapconsist{\Omega}$ and 
From Lemma~\ref{lemma:omegaclosure}, we have $mapconsist{\Omega^*}$, so there
must be some $r_j$ such that $M(r_i') \subseteq M(r_j)$.  Therefore $v \in M(r_j)$
and $\valueconsist{v}{T'@(\oton{r})}$.

\item Case [E-DownRgn].
By assumption, $$\opsenv{\text{downregion}(e,\oton{r})}{v}{E}$$ and
$$\typeenv{\text{downregion}(e,\oton{r})}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The type rule must be [T-DownRgn], so we also know $T = T' @ (\oton{r})$.

@@

To show $\valueconsist{v}{T}$ there are two cases.  If $l \not\in C$, then $v = S'(l)$.
Since $\storeconsist{S'}{H}$ from Lemma~\ref{lemma:heapconst:apply}, we have $\valueconsist[]{S'(l)}{H(l)}$.
If $l \in C$, we have $v = \valueconsist[]{v'}{H(l)}$ from [E-Read] directly.  In
both cases, $\valueconsist{l}{T@(\oton{r})}$ gives us $H(l) = \hat T$ and therefore
$\valueconsist{v}{T}$.
%If $v = M(l)$ then the claim follows from $\valueconsist{l}{T@(\oton{r})}$.
%If $v$ is instead some $v'$ of type $H(l)$ the result also follows because $H(l) = \hat{T}$.
Next, $\traceconsist{E \dplus [ read(l,excl,v)]}$ follows immediately from $\traceconsist{E}$ as
any read operation is consistent (the restriction is only on writes and reductions).

@@

Observe that 
$$M \llbracket \exists r'_1, \ldots r'_n.\ T_2\text{ where }\Omega_1 \rrbracket$$
replaces all free region names in $T_2$ and particularly $\Omega_1$ by physical regions, so
that the only region names remaining are $r'_1,\ldots,r'_n$.  It is then easy to show that
$$\mapconsist[\{(r'_i,\rho_i)\}]{M \backslash \{r'_1,\ldots, r'_n \}\llbracket \Omega_1 \rrbracket}$$
from which $\valueconsist{v'}{T_1}$ follows (see Figure~\ref{fig:constprop}).

@@

To show $\localconsist[M']{L'}{\Gamma'}$, we must show
$\valueconsist[M']{L'(x)}{\Gamma'(x)}$ for all identifiers $x$ in
the domain of $\Gamma'$.  For any $id' \neq id$, the statement is equivalent
to showing $\valueconsist[M']{L(x)}{\Gamma(x)}$, which must hold because $M'$ does 
does not change any existing binding in $M$.  For $id$, we have
$\valueconsist[M']{v_1}{T_1[r_1/r'_1,\ldots,r_k/r'_k]}$, which is implied by
$\valueconsist{\langle \langle \otok{k}{\rho},v \rangle\rangle}{T_1}$.

@@

To show that
$\traceconsist{E_{n+1}}$ under $M'$ implies $\traceconsist{E_{n+1}}$
under $M$, note that $H(\cdot)$ does not depend on the mapping (because it includes only physical, not logical, regions) so the
types of locations are unaffected by the mapping (see
Figure~\ref{fig:constprop}).  Now, the environment
$M' \llbracket \Gamma' \rrbracket$ is used to look up reduction
functions, but the reduction function signatures are restricted to
have no regions named in their arguments and so
$$M \llbracket \Gamma' \rrbracket(id)  = M' \llbracket \Gamma' \rrbracket(id)$$
Finally, from 
$$valid\_interleave(S, C, E'', E', E_{n+1})$$
and Lemma~\ref{lemma:heapconst:effects2} we conclude $\traceconsist{E''}$.

For the last consistency property of the function call, we first observe
that $\privconsist[M']{E_{n+1}}{\Phi'}$ is equivalent to 
$$\privconsist{E_{n+1}}{\Phi[r_1/r'_1,\ldots,r_k/r'_k]}$$
because, again, $E_{n+1}$ does not mention logical region names and
$M' \llbracket r' \rrbracket = M \llbracket r \rrbracket$.
From [T-Call] we have 
$$\Phi[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Phi^*$$
and so
$\privconsist{E_{n+1}}{\Phi^*}$.  Applying Lemma~\ref{lemma:phiclosure} we conclude
that $\privconsist{E_{n+1}}{\Phi}$.
Finally, from Lemma~\ref{lemma:effsound:effects2} we conclude $\privconsist{E''}{\Phi}$.
\end{itemize}}
\end{proof}
